Global Optimization of the Analog Method by Means of Genetic Algorithms

### 2016-08-17 Genetic algorithms - Journal of Climate/template.tex
Changing "analogue" for "analog".

\title{Global Optimization of the Analogue Method by Means of Genetic Algorithms}

@@

\abstract{The Analogue Method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic sequential procedure. This calibration approach has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely Genetic Algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. Multiple combinations of algorithms had to be assessed, and even new operators were developed, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of Genetic Algorithms to optimize the Analogue Method. These recommendations are the main outcome of this work. It opens new perspective for the improvement of the Analogue Method, and its application to new regions or to new predictands.
}

@@

The Analogue Method (AM) relies on the hypothesis that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. The principle consists in sampling a certain number of past situations based on different atmospheric variables (predictors) in order to build a probabilistic forecast for a local weather variable of interest (predictand). The most common usage of the AM is for precipitation forecasting \citep[eg.][]{Guilbaud1997, Bontron2005, Bliefernicht2010, Marty2012, Horton2012, Radanovics2013, BenDaoud2015}, but AMs, or equivalent, were also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. Applications for monthly forecasts in many countries also exist, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

Even though the method is rather simple compared to numerical weather prediction (NWP) models, it contains a certain number of parameters one needs to determine, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criteria itself, and finally the number of analogue situations to keep at each subsampling level.

@@

Up to now, the optimization of the method has been undertaken by semi-automatic sequential calibration procedures \citep[see][for its definition]{Bontron2004, Horton2016}. However, the selection of predictor variables, pressure levels and temporal windows had still to be made manually before optimizing the spatial windows and the number of analogues. Testing multiple combinations of these is very combinatorial and becomes quickly cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing the method with the sequential technique is laborious, as many combinations of predictors (variables, pressure levels, temporal windows) have to be assessed. Moreover, proceeding to the optimization sequentially ignores potential dependencies between the parameters of the method, may they be within a single level of analogy or between them, which could lead to another parametrization if the parameters were calibrated together. Simultaneous calibration of all parameters has never been undertaken so far. Thus, due to the sequential approach, the risk of ending in a local optimum is high and can not be avoided. Indeed, during the AM calibration it has been found that the resulting parameters may vary with initial choices (such as the number of analogues).

@@

\section{The Analogue Method}
\label{sec:am}

@@

The method based on the analogy of the synoptic circulation consists in the following steps: the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criteria \citep{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, on 2 different the geopotential heights (e.g. 500 and 1000~hPa) and over a certain spatial window. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centered around the target date, for every year of the archive. The $n_{1}$ (parameter to calibrate) dates with the lowest values of S1 are considered as analogues to the target day. Then, the daily observed precipitation amount of the $n_{1}$ resulting dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day.

The other current parametrization adds a second level of analogy on moisture variables, more specifically on a moisture index made of the product of the precipitable water with the relative humidity at a certain pressure level \citep[e.g. 850~hPa, see][]{Bontron2004}. When adding a second level of analogy, $n_{2}$ dates are subsampled in the $n_{1}$ analogues on the atmospheric circulation, to end up with a smaller number of analogue situations.

The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $l$ can be written:

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (set of parameters of the AM in this case). Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g. geopotential, or temperature), discrete (e.g. number of analogues to select, from 1 to 40), or continuous. For example, a chromosome may looks like:

\begin{equation}
[p_{i,j}, h_{i,j}, \lambda w_{i,j}, \lambda e_{i,j}, \varphi s_{i,j}, \varphi n_{i,j}, p_{i,j+1}, h_{i,j+1}, ..., n_{i}, ... p_{i+1,j}, h_{i+1,j}, ..., n_{i+1}, ...]
\end{equation}
where, for the level of analogy $i$ and the predictor number $j$, $p_{i,j}$ is the pressure level, $h_{i,j}$ is the temporal window, $\lambda w_{i,j}$ and $\lambda e_{i,j}$ are the western and eastern longitudes of the spatial window, $\varphi s_{i,j}$ and $\varphi n_{i,j}$ are its southerly and northerly latitudes, and $n_{i,j}$ is the number of analogues. A chromosome for the simple AM on the atmospheric circulation only (Sect. \ref{sec:am}) may be: $[500~hPa, 12~h, -2.5\degree, 15\degree, 40\degree, 55\degree, 1000~hPa, 24~h, 0\degree, 15\degree, 42.5\degree, 47.5\degree, 28]$

>>>>

\title{Global Optimization of the Analog Method by Means of Genetic Algorithms}

@@

\abstract{The Analog Method is based on a statistical relationship between synoptic atmospheric variables (predictors) and local weather (predictand), which we aim at forecasting. This relationship is expressed through many parameters that are usually calibrated by means of a semi-automatic sequential procedure. This calibration approach has strong limitations: it is made of successive steps and thus cannot handle parameters dependencies, and it cannot automatically optimize some parameters, such as the selection of the pressure levels and the temporal windows on which the predictors are compared.
In order to surpass these limitations, a global optimization technique was assessed, namely Genetic Algorithms, which can optimize jointly all parameters of the method and get closer to a global optimum by taking into account the parameters dependencies. Moreover, it can choose objectively parameters that were previously manually assessed, and can take into account new degrees of freedom that were unthinkable before.
These kind of optimization techniques need however to be tailored to the problem to solve. Multiple combinations of algorithms had to be assessed, and even new operators were developed, such as the \textit{chromosome of adaptive search radius} which was found to be very robust, in order to make recommendations for the use of Genetic Algorithms to optimize the Analog Method. These recommendations are the main outcome of this work. It opens new perspective for the improvement of the Analog Method, and its application to new regions or to new predictands.
}

@@

The Analog Method (AM) relies on the hypothesis that similar situations in terms of atmospheric circulation are likely to lead to similar local weather \citep{Lorenz1956, Lorenz1969, Duband1970, Bontron2005}. The principle consists in sampling a certain number of past situations based on different atmospheric variables (predictors) in order to build a probabilistic forecast for a local weather variable of interest (predictand). The most common usage of the AM is for precipitation forecasting \citep[eg.][]{Guilbaud1997, Bontron2005, Bliefernicht2010, Marty2012, Horton2012, Radanovics2013, BenDaoud2015}, but AMs, or equivalent, were also used for short to medium term forecasting of daily temperatures \citep{Radinovic1975, Woodcock1980, Kruizinga1983}, wind \citep{Gordon1987}, snow avalanches \citep{Obled1980, Bolognesi1993}, insolation \citep{Bois1981}, and the trajectory of tropical cyclones \citep{Keenan1981, Sievers2000, Fraedrich2003}. Applications for monthly forecasts in many countries also exist, including Canada \citep{Shabbar1986},  Hungary \citep{Toth1989}, the Netherlands \citep{Nap1981}, and England \citep{Murray1974}, as well as seasonal forecasts: \citet{Barnett1978}, \citet{Bergen1982} and \citet{Livezey1988}.

Even though the method is rather simple compared to numerical weather prediction (NWP) models, it contains a certain number of parameters one needs to determine, such as the choice of the predictor variable, its pressure level and temporal window to consider, the spatial domain to use for the comparison, as well as the analogy criteria itself, and finally the number of analog situations to keep at each subsampling level.

@@

Up to now, the optimization of the method has been undertaken by semi-automatic sequential calibration procedures \citep[see][for its definition]{Bontron2004, Horton2016}. However, the selection of predictor variables, pressure levels and temporal windows had still to be made manually before optimizing the spatial windows and the number of analogs. Testing multiple combinations of these is very combinatorial and becomes quickly cumbersome, especially when considering multiple predictors within the same level of analogy. Thus, optimizing the method with the sequential technique is laborious, as many combinations of predictors (variables, pressure levels, temporal windows) have to be assessed. Moreover, proceeding to the optimization sequentially ignores potential dependencies between the parameters of the method, may they be within a single level of analogy or between them, which could lead to another parametrization if the parameters were calibrated together. Simultaneous calibration of all parameters has never been undertaken so far. Thus, due to the sequential approach, the risk of ending in a local optimum is high and can not be avoided. Indeed, during the AM calibration it has been found that the resulting parameters may vary with initial choices (such as the number of analogs).

@@

\section{The Analog Method}
\label{sec:am}

@@

The method based on the analogy of the synoptic circulation consists in the following steps: the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criteria \citep{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, on 2 different the geopotential heights (e.g. 500 and 1000~hPa) and over a certain spatial window. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centered around the target date, for every year of the archive. The $n_{1}$ (parameter to calibrate) dates with the lowest values of S1 are considered as analogs to the target day. Then, the daily observed precipitation amount of the $n_{1}$ resulting dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day.

The other current parametrization adds a second level of analogy on moisture variables, more specifically on a moisture index made of the product of the precipitable water with the relative humidity at a certain pressure level \citep[e.g. 850~hPa, see][]{Bontron2004}. When adding a second level of analogy, $n_{2}$ dates are subsampled in the $n_{1}$ analogs on the atmospheric circulation, to end up with a smaller number of analog situations.

The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analog situations, compared to the observed value $y^{0}$. The better the prediction, the smaller the score. The mean CRPS of a prediction series of length $l$ can be written:

@@

The GAs optimize a population of $N$ individuals (parameter sets). Each individual contains a chromosome (set of parameters of the AM in this case). Genes are the individual parameters constituting the chromosome. They can be either categorical (e.g. geopotential, or temperature), discrete (e.g. number of analogs to select, from 1 to 40), or continuous. For example, a chromosome may looks like:

\begin{equation}
[p_{i,j}, h_{i,j}, \lambda w_{i,j}, \lambda e_{i,j}, \varphi s_{i,j}, \varphi n_{i,j}, p_{i,j+1}, h_{i,j+1}, ..., n_{i}, ... p_{i+1,j}, h_{i+1,j}, ..., n_{i+1}, ...]
\end{equation}
where, for the level of analogy $i$ and the predictor number $j$, $p_{i,j}$ is the pressure level, $h_{i,j}$ is the temporal window, $\lambda w_{i,j}$ and $\lambda e_{i,j}$ are the western and eastern longitudes of the spatial window, $\varphi s_{i,j}$ and $\varphi n_{i,j}$ are its southerly and northerly latitudes, and $n_{i,j}$ is the number of analogs. A chromosome for the simple AM on the atmospheric circulation only (Sect. \ref{sec:am}) may be: $[500~hPa, 12~h, -2.5\degree, 15\degree, 40\degree, 55\degree, 1000~hPa, 24~h, 0\degree, 15\degree, 42.5\degree, 47.5\degree, 28]$
